---
title: "Olympic Medals 2024"
format: html
execute:
  warning: false
  echo: false
editor_options: 
  chunk_output_type: console
fig-width: 10
fig-asp: 0.618
out-width: "100%"
fig-align: center
---

```{r}
#| label: setup

# Load libraries for functions to work
library(httr)
library(rvest)
library(tidyverse)
library(magrittr)
library(purrr)
library(janitor)
library(ggrepel)

# Source functions
source("R/Scraping Functions.R")

# Define plot style
colors <- c("Gold" = "gold", "Silver" = "#C0C0C0", "Bronze" = "#CD7F32")
linecolor <- c("darkorange", "cadetblue")
plot_style <- theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 20, face = "bold"),
        panel.grid.minor.x = element_blank(),
        strip.text = element_text(size = 20, face = "bold"),
        axis.text = element_text(size = 14),
        axis.title = element_text(size = 14),
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 14))

```

### What's in here?

In this post, we'll compare the officially communicated goals that nations are aiming for against the reality of Paris 2024 Olympic Games.

Stay tuned and visit daily for the latest updates, as this report will be refreshed every day to keep you informed.

*Last update date: `r paste0(format(Sys.time(), format = "%F %R"), "h")`*

### Select your Country

::: panel-tabset
## Spain

Spain is dreaming big for the Paris 2024 Olympic Games.

No one is hiding that [the medal target](https://elpais.com/deportes/juegos-olimpicos/2024-07-08/la-delegacion-espanola-para-los-juegos-olimpicos-mas-deportistas-mas-mujeres-y-mas-equipos.html){target="_blank"} is to beat the 22 medals won in Barcelona’s 1992 Olympics, the best Olympic Games of the country's history.

But is Spain on the right track?

```{r}
#| label: load data spain

# Although functions work with more values, I recommend to only load 1 country and 2 years for comparison
selected_year <- c(1992, 2024)
selected_country <- c("Spain")

olympics_medal_table <- scrape_and_clean_olympics_wiki_multiple(selected_year = selected_year, selected_country = selected_country)

```

```{r}
#| label: prepare data spain

olympics_list <- transform_olympics_data_as_list(olympics_medal_table)

```

```{r}
#| label: define plot style spain

linetypes <- setNames(c("dotted", "solid"), selected_year) # use setNames() to enquote numeric values

```

```{r}
#| label: plot running total comparison spain

plot <- plot_olympics_data(olympics_list$olympics_medal_table_summarized_pivoted,
                           olympics_list$max_days,
                           selected_country[[1]])
plot +
    labs(x = "Day",
         y = "Running Total of Medals",
         color = "Year") +
    scale_x_continuous(breaks = 0 : max(olympics_list$max_days$max_day)) +
    scale_color_manual(values = linecolor) +
    scale_linetype_manual(values = linetypes, guide = "none") +
  plot_style

```

```{r}
#| label: fitted model spain

fitted_model <- add_slope_olympics_data(olympics_list$olympics_medal_table_summarized_pivoted)

```

### Gold, Silver or Bronze?

To reach the `r fitted_model %>% filter(year == selected_year[[1]]) %>% pull(max_running_total_year_country)` medals they won in `r selected_year[[1]]`, they now need to win an average of `r fitted_model$slope[[1]] %>% round(1)` medals per day.

The current count of silver and bronze medals is equal or higher than that achieved in Barcelona, but it's the gold medals that are the missing. Can Spain make up for its gold medal shortcomings?

```{r}
#| label: plot running total per medal comparison spain
#| fig-height: 8

plot <- plot_olympics_data_per_medal(olympics_list$olympics_medal_table_summarized,
                                     selected_country[[1]])

plot +
  labs(x = "Day",
       y = "Running Total of Medals",
       linetype = "Year") +
  scale_x_continuous(breaks = 0 : max(olympics_list$max_days$max_day)) +
  scale_color_manual(values = colors, guide = "none") +
  scale_linetype_manual(values = linetypes) +
  plot_style

```

### Milestones

With 382 athletes, Spain presents the second-largest contingent in their history of the Games (the record number of 430 was in Barcelona '92). And for the first time, there are more women (192) than men (190) in the delegation.

We still have a week to go and I will keep you updated here!

## Germany

In a first step, Germany hopes to stop the downward trend in the medals table.

[The official goal](https://www.sueddeutsche.de/sport/olympia-deutschland-medaillenspiegel-ziel-lux.EWYjh2AYTmCducSPwTBPi2){target="_blank"} is to stay in the top 10 like three years ago in Tokyo and to bounce back from that worst result since their reunification.

But is Germany on the right track?

```{r}
#| label: load data germany

# Although functions work with more values, I recommend to only load 1 country and 2 years for comparison
selected_year <- c(2020, 2024)
selected_country <- c("Germany")

olympics_medal_table <- scrape_and_clean_olympics_wiki_multiple(selected_year = selected_year, selected_country = selected_country)

```

```{r}
#| label: prepare data germany

olympics_list <- transform_olympics_data_as_list(olympics_medal_table)

```

```{r}
#| label: define plot style germany

linetypes <- setNames(c("dotted", "solid"), selected_year) # use setNames() to enquote numeric values

```

```{r}
#| label: plot running total comparison germany

plot <- plot_olympics_data(olympics_list$olympics_medal_table_summarized_pivoted,
                           olympics_list$max_days,
                           selected_country[[1]])
plot +
    labs(x = "Day",
         y = "Running Total of Medals",
         color = "Year") +
    scale_x_continuous(breaks = 0 : max(olympics_list$max_days$max_day)) +
    scale_color_manual(values = linecolor) +
    scale_linetype_manual(values = linetypes, guide = "none") +
  plot_style

```

```{r}
#| label: fitted model germany

fitted_model <- add_slope_olympics_data(olympics_list$olympics_medal_table_summarized_pivoted)

```

### Gold, Silver or Bronze?

They need to win `r fitted_model$slope[[1]] %>% round(1)` medals on average per day to match their `r fitted_model %>% filter(year == selected_year[[1]]) %>% pull(max_running_total_year_country) %>% round(1)` medals from `r selected_year[[1]]`.

While the total medal target seems far away, with the number of gold medals they've already won, Germany is well on their way to matching their last position.

Are the going to reach the 10 times gold from Tokyo?

```{r}
#| label: plot running total per medal comparison germany
#| fig-height: 8

plot <- plot_olympics_data_per_medal(olympics_list$olympics_medal_table_summarized,
                                     selected_country[[1]])

plot +
  labs(x = "Day",
       y = "Running Total of Medals",
       linetype = "Year") +
  scale_x_continuous(breaks = 0 : max(olympics_list$max_days$max_day)) +
  scale_color_manual(values = colors, guide = "none") +
  scale_linetype_manual(values = linetypes) +
  plot_style

```

## France

France's Olympic dream is evolving from host to champion.

[Their medal mission](https://www.lemonde.fr/sport/article/2021/10/14/jo-2024-de-grandes-ambitions-mais-des-petites-foulees-pour-le-sport-francais_6098291_3242.html){target="_blank"}, officially communicated by Emmanuel Macron, is to return to the top 5 as in Atlanta 1996, where they have so far obtained their gold medal record.

But is France on the right track?

```{r}
#| label: load data france

# Although functions work with more values, I recommend to only load 1 country and 2 years for comparison
selected_year <- c(1996, 2024)
selected_country <- c("France")

olympics_medal_table <- scrape_and_clean_olympics_wiki_multiple(selected_year = selected_year, selected_country = selected_country)

```

```{r}
#| label: prepare data france

# Manually fill missing values from Wikipedia table :-(
olympics_medal_table_imputed <- olympics_medal_table %>% 
  mutate(date_cod = 
           case_when(
             medal == "Gold" &
               name == "Christophe CapellePhilippe ErmenaultJean-Michel MoninFrancis Moreau" &
               sport == "Cycling" &
               event == "Men's team pursuit" ~ as_date("1996-07-27"),
             medal == "Gold" &
               name == "Nathalie Lancien" &
               sport == "Cycling" &
               event == "Women's points race" ~ as_date("1996-07-28"),
             medal == "Gold" &
               name == "Jean-Pierre Amat" &
               sport == "Shooting" &
               event == "Men's 50 m rifle three positions" ~ as_date("1996-07-27"),
             medal == "Silver" &
               name == "Philippe Ermenault" &
               sport == "Cycling" &
               event == "Men's individual pursuit" ~ as_date("1996-07-26"),
             medal == "Silver" &
               name == "Marion Clignet" &
               sport == "Cycling" &
               event == "Women's individual pursuit" ~ as_date("1996-07-28"),
             medal == "Silver" &
               name == "Jeannie Longo" &
               sport == "Cycling" &
               event == "Women's road time trial" ~ as_date("1996-08-03"),
             medal == "Bronze" &
               name == "Jean-Pierre Amat" &
               sport == "Shooting" &
               event == "Men's 10 m air rifle" ~ as_date("1996-07-22"),
             .default = date_cod))

olympics_list <- transform_olympics_data_as_list(olympics_medal_table_imputed)

```

```{r}
#| label: define plot style france

linetypes <- setNames(c("dotted", "solid"), selected_year) # use setNames() to enquote numeric values

```

```{r}
#| label: plot running total comparison france

plot <- plot_olympics_data(olympics_list$olympics_medal_table_summarized_pivoted,
                           olympics_list$max_days,
                           selected_country[[1]])
plot +
    labs(x = "Day",
         y = "Running Total of Medals",
         color = "Year") +
    scale_x_continuous(breaks = 0 : max(olympics_list$max_days$max_day)) +
    scale_color_manual(values = linecolor) +
    scale_linetype_manual(values = linetypes, guide = "none") +
  plot_style

```

```{r}
#| label: fitted model france

fitted_model <- add_slope_olympics_data(olympics_list$olympics_medal_table_summarized_pivoted)

```

### Gold, Silver or Bronze?

The `r selected_year[[1]]` team has set the bar high with `r fitted_model %>% filter(year == selected_year[[1]]) %>% pull(max_running_total_year_country) %>% round(1)` medals, and they're aiming to reach that number once again. To achieve this, they'll need to win `r fitted_model$slope[[1]] %>% round(1)` medals on average per day.

Can they set a new record?

```{r}
#| label: plot running total per medal comparison france
#| fig-height: 8

plot <- plot_olympics_data_per_medal(olympics_list$olympics_medal_table_summarized,
                                     selected_country[[1]],
                                     with_labels = FALSE)

plot +
  labs(x = "Day",
       y = "Running Total of Medals",
       linetype = "Year") +
  scale_x_continuous(breaks = 0 : max(olympics_list$max_days$max_day)) +
  scale_color_manual(values = colors, guide = "none") +
  scale_linetype_manual(values = linetypes) +
  plot_style

```
:::

### References

-   Data are downloaded from Wikipedia and depend on the source being up to date.
-   [Olympic Games overview](https://en.wikipedia.org/wiki/List_of_Olympic_Games_host_cities){target="_blank"}.
-   Spain: [1992](https://en.wikipedia.org/wiki/Spain_at_the_1992_Summer_Olympics){target="_blank"} and [current performance](https://en.wikipedia.org/wiki/Spain_at_the_2024_Summer_Olympics){target="_blank"}.
-   Germany: [2020](https://en.wikipedia.org/wiki/Germany_at_the_2020_Summer_Olympics){target="_blank"} and [current performance](https://en.wikipedia.org/wiki/Germany_at_the_2024_Summer_Olympics){target="_blank"}.
-   France: [1996](https://en.wikipedia.org/wiki/France_at_the_1996_Summer_Olympics){target="_blank"} and [current performance](https://en.wikipedia.org/wiki/France_at_the_2024_Summer_Olympics){target="_blank"}.
-   Check out my [GitHub repository](https://github.com/dani-f){target="_blank"} to see how this report was compiled and adapt it to your country.
